---
title: "Coursework 1"
author: "Patrick Matthews"
date: "2023-11-21"
output: html_document
---
Question R1
Define the vector x <- c(5,9,2,3,4,6,7,0,8,12,2,9). For each of the following questions, you should use just a single line of code in R to calculate the answer (10 marks):

a) What is the sum of the first 4 elements? 
```{r}
x <- c(5,9,2,3,4,6,7,0,8,12,2,9)
sum (x [1:4])
```
After defining the vector x <- c(5,9,2,3,4,6,7,0,8,12,2,9) the sum command was used to add up the values in the vector. This command was paired with x [1:4] which tells the programme to add up elements one to four in the vector x giving the answer 19. 

b) What is the highest value in the vector? 
```{r}
max (x)
```
The max command outputs the highest value in a vector which is defined. In this case it searches through the vector x and outputs the highest value, in this case it as 12. 

c ) Which position in the vector contains the highest value? 
```{r}
which.max(x)
```
The which.max command is similar to the max command but instead of outputting the value of the largest number it outputs the position. In this case the largest value ,12, in vector x is in the tenth position. 

d) Which position in the vector would the highest value occur if the vector is reversed in order?
```{r}
which.max(rev(x))
```
This combines commands, being able to put the rev command within the which.max and the rev command simply completely reverses the vector x so it is now back to front with the new position of the largest value ,12, being in position three. 

e) How many values in the vector are lower than 10, but higher than 5? 
```{r}
x > 5 & x < 10
```
This code utilises the greater than (>) and the less than (<) signs to give the values in x between six and nine. This outputs as Boolean being true or false and if count all true’s being in this inequality there are five numbers in the vector x between six and nine. 

Question R2
For this question, use basic R commands. The file ‘cell_features.csv’ contains information from different cell lines and experiments that have been run in different fibronectin concentrations. Create a new data structure containing cell area means from batch A only and then calculate the median value and interquartile range of these. Now create a new vector of these cell area means where outliers are removed (defined as values 2 times the IQR above or below the median value) (14 marks).
```{r}
read.csv ("cell_features")
new_data <- cell_features[cell_features$batch == "Batch A", "cell_area_mean"]
```
This creates a new data structure called new_data which the specific data from the cell_features file can be extracted into. In this case it is asking to console to look into the cell_features file specifically in the batch column and take out every string in the batch column with “Batch A” and pair this with their corresponding cell_area_mean value.
